AggregateCounts

### 2021-06-15 R/AggregateCounts.Rmd
added new histogram function

```{r setup, include=FALSE}
library(conflicted)
library(tidyverse)
filter <- dplyr::filter

ReadAggregateFile(fileList[1], "Pa", "1", "wt")
blah<-GetAggregates()

datos <-read.csv(fileList[1], header = F)
colnames(datos)<-c('Size')
datos1<-datos %>% filter(!Size<0.5) #first filter the size

ReadAggregateFile<-function(file, species, timepoint, condition, replicate){#make a dataframe of the aggregate size list and add experimental parameters
   datos <-read.csv(file, header = F)
   colnames(datos)<-c('Size')
   datos<-datos %>% filter(!Size<0.5) #first filter the size
   datos<-datos %>% mutate(Species = species)#add parameters
   datos<-datos %>% mutate(TimePoint = timepoint)
   datos<-datos %>% mutate(Condition = condition)
   datos<-datos %>% mutate(Replicate = replicate)
   datos$Species<-as.factor(datos$Species)
   datos$Replicate<-as.factor(datos$Replicate)
   datos$Condition<-as.factor(datos$Condition)
   datos$TimePoint<-as.factor(datos$TimePoint)
   return(datos)
}

GetAggregates<-function(){#import all experiments at once
   folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Aggregate lists')
   fileList<-list.files(path = folder, full.names = TRUE)
   growthDf<-data.frame()
   for (aggList in 1:length(fileList)){
     fileName<-str_split(basename(fileList[aggList]), "\\.")[[1]][1]#this selects the base file name and removes the ".csv"
     attrList<-str_split(fileName,"_")[[1]]
     growthDf<-rbind(growthDf,ReadAggregateFile(fileList[aggList],attrList[4],attrList[2],attrList[1],attrList[3]))
     }
   return(growthDf)
}
```

>>>>

```{r setup, include=FALSE}
library(conflicted)
library(tidyverse)
library(plotrix)#for std error
library(dplyr)
library(cowplot)
filter <- dplyr::filter
source("~/Documents/Whiteley/PROJECTS/SaPa Chronic Wounds/PaSaChronicWounds/R/readTiff.R")
library(EBImage) #for rtiff

aggregateList<-GetAggregates()#this reads the data from Matlab
# mutate(Strain = recode(Position, "0" = "Pa14 wt", "1"= "Pa14 -PA1414","2" = "Pa14 wt", "3"= "Pa14 -PA1414")) #this is to change the names

#agregate input functions
#####
ReadAggregateFile<-function(file, species, timepoint, condition, replicate){#make a dataframe of the aggregate size list and add experimental parameters 
   datos <-read.csv(file, header = F)  
   colnames(datos)<-c('Size') 
   datos<-datos %>% filter(!Size<0.5) #first filter the size
   datos<-datos %>% mutate(Species = species)#add parameters
   datos<-datos %>% mutate(TimePoint = timepoint) 
   datos<-datos %>% mutate(Condition = condition)
   datos<-datos %>% mutate(Replicate = replicate)
   datos$Species<-as.factor(datos$Species)
   datos$Replicate<-as.factor(datos$Replicate)
   datos$Condition<-as.factor(datos$Condition)
   datos$TimePoint<-as.factor(datos$TimePoint)
   return(datos)
}

GetAggregates<-function(){#import all experiments at once
   folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Aggregate lists')
   fileList<-list.files(path = folder, full.names = TRUE)
   growthDf<-data.frame()
   for (aggList in 1:length(fileList)){
     fileName<-str_split(basename(fileList[aggList]), "\\.")[[1]][1]#this selects the base file name and removes the ".csv"
     attrList<-str_split(fileName,"_")[[1]]
     growthDf<-rbind(growthDf,ReadAggregateFile(fileList[aggList],attrList[4],attrList[2],attrList[1],attrList[3]))
     }
   return(growthDf)
}
#####
```
```{r aggregate wrangling}
#Statistics from all aggregates

#Get aggregate summary per sample
aggregateListSummaryReplicate<-aggregateList %>% group_by(Species, TimePoint, Condition, Replicate) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass )

#Plot agg summary per sample
#####
ggplot()+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= meanAggSize, color = Species,shape = "MeanSize"),size = 3, height = 0, width = .1)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= maxSize, color = Species, shape = "MaxSize"),size = 3, height = 0, width = .1)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= biomass, color = Species, shape = "TotalBiomass"),size = 3, height = 0, width = .1)+  
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(. ~ Condition, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per sample')+
  labs(x = "",y = "um3")+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5))
#####

#Get aggregate Summary per condition
aggregateListSummary<-aggregateList %>% group_by(Species, TimePoint, Condition) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass ) 
#plot agg summary per condition
#####
ggplot()+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= meanAggSize, color = Species,shape = "MeanSize"),size = 3, height = 0, width = .1)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= maxSize, color = Species, shape = "MaxSize"),size = 3, height = 0, width = .1)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= biomass, color = Species, shape = "TotalBiomass"),size = 3, height = 0, width = .1)+  
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(. ~ Condition, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per sample')+
  labs(x = "",y = "um3")+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5))
#####
```
```{r non-fancy histogram}
#histogram<- hist(aggregateList$Size, breaks = c(0, 5, 10, 100, 1000, 10^9))#this is a simple way to do a histogram

#this pipe essentially gets to the producto of what fancy histograms do but more efficiently
#First including all sizes
#####
newHistogram.df<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% group_by(Species,Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% group_by(Species,Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

#This is weird, but I want to have the histogram above with all the data, and the one below with only the bins
newHistogramSummary.df<-newHistogram.df %>% group_by(Species,Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#now plot it
ggplot()+
    geom_jitter(data = newHistogramSummary.df,aes(x=as.factor(LogBin),y=NormalizedFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))  
#####
#now only aggregates 5 microns or larger
#####
aggsOnlyHist.df<-aggregateList %>% mutate(LogBin = case_when(Size <= 5 ~ 5,
                        Size<= 10 ~ 10,
                        Size<= 100    ~ 100,
                        Size<= 1000   ~ 1000,
                        Size<= 10000  ~ 10000,
                        Size<= 100000 ~ 100000,
                        Size > 100000 ~ 1000000)) %>% filter(!LogBin == 5) %>% group_by(Species,Condition,TimePoint, Replicate)%>% 
  mutate(TotalBiomass = sum(Size), TotalFrequency = n()) %>% ungroup() %>% group_by(Species,Condition,TimePoint, Replicate, LogBin)%>%
  mutate(BinBiomass = sum(Size), BinFrequency = n()) %>% ungroup() %>%
  mutate(NormalizedFrequency = BinFrequency/TotalFrequency*100, NormalizedBiomass = BinBiomass/TotalBiomass*100)

#This is weird, but I want to have the histogram above with all the data, and the one below with only the bins
aggsOnlyHistSummary.df<-aggsOnlyHist.df %>% group_by(Species,Condition,TimePoint, Replicate, LogBin,NormalizedFrequency,NormalizedBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkSumFreq =sum(NormalizedFrequency), checkSumBiomass = sum(NormalizedBiomass))

#now plot it
#normalized frequency
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet timepoint, color species
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass,  color =Species, ),size=3, height = 0, width = .2)+
    scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet species, color timepoint
ggplot()+
    geom_jitter(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass,  color =TimePoint, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~Species)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
#Normalized Biomass facet species and Timepoint, color Replicate
ggplot()+
    geom_line(data = aggsOnlyHistSummary.df,aes(x=as.factor(LogBin),y=NormalizedBiomass, color = Replicate, group = Replicate ),size=1)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    guides(color = FALSE)+
    facet_grid(TimePoint~Species)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    scale_colour_grey()+
    theme(plot.title = element_text(hjust = 0.5))

#####
#Planktonic vs aggregates
#####
#Add agregation value to the dataframe
newHistogram.df<-newHistogram.df %>%  mutate(Aggregation = case_when(LogBin == 5 ~ "Planktonic",LogBin>5 ~ "Aggregate")) %>%
  group_by(Species,Condition,TimePoint, Replicate, Aggregation)%>% 
  mutate(AgPlBiomass = sum(Size), AgPlFreq = n()) %>% ungroup() %>%
  mutate(AgPlNormFrequency = AgPlFreq/TotalFrequency*100, AgPlNormBiomass = AgPlBiomass/TotalBiomass*100)
#Sumarize in dataframe
plankAggHist<-newHistogram.df %>% group_by(Species,Condition,TimePoint, Replicate, Aggregation,AgPlNormFrequency,AgPlNormBiomass) %>% summarize() %>%  unique()%>%ungroup() %>%  group_by(Species,Condition,TimePoint, Replicate) %>% mutate(checkAggFreq =sum(AgPlNormFrequency), checkAggBiomass = sum(AgPlNormBiomass))
#plot Frequency
ggplot()+
    geom_jitter(data = plankAggHist,aes(x=as.factor(Aggregation),y=AgPlNormFrequency,  color =Species, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5)) 
ggplot()+
    geom_jitter(data = plankAggHist,aes(x=as.factor(Aggregation),y=AgPlNormBiomass,  color =Species, ),size=3, height = 0, width = .2)+
    #scale_y_log10(breaks = c(100, 10, 1, .1, .01), labels = c(100, 10, 1, .1, .01))+
    facet_grid(.~TimePoint)+
    labs(y = "Normalized Biomass",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

#####
```

```{r fancy histogram of aggregates}
#Create histograms
##this takes a long time, so consider running them in parts
allDataGrouped<-aggregateList %>% group_by(Species, TimePoint, Condition)#first do the correct grouping
#allDataGrouped<-aggregateList %>% group_by(Species, TimePoint, Condition, Sample)# this is to include sample on the distribution or keep it out
allDataList<- group_split(allDataGrouped)#create list of dataframes
allDataHistograms.5<-map2(allDataList, 5,TotalFancyHistogram)#apply histogram function to all items in the list
allDataHistograms.5df<- bind_rows(allDataHistograms.5)#bring them together

#plotting histogram data
hist.5.LogBinSummarize<-allDataHistograms.5df %>% 
  filter(!max == 5)%>% #remove planktonic cells, but this changes the nortmalization of aggregate counts, so the normalization needs to be done later for aggs only
  group_by(Species,Condition,TimePoint)%>% mutate(LogBin = case_when(max <= 10 ~ 10,
                        max<= 100    ~ 100,
                        max<= 1000   ~ 1000,
                        max<= 10000  ~ 10000,
                        max<= 100000 ~ 100000,
                        max > 100000 ~ 1000000)) %>% ungroup() %>% 
  group_by(Species, Condition,TimePoint,LogBin) %>% summarise(FrequencyLogBin = sum(frequency), BiomassLogBin = sum(BiomassBin), DensityLogBin = sum(Density))

#plot frequency  
ggplot()+
    geom_point(data = hist.5.LogBinSummarize,aes(x=as.factor(LogBin),y=FrequencyLogBin,  color =Species ),size=4)+
    #labs(title = titulo)+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~TimePoint)+
    labs(y = "Frequency",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    #guides(color = FALSE)+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

  ggplot()+
    geom_point(data = hist.5.LogBinSummarize,aes(x=as.factor(LogBin),y=BiomassLogBin,  color =Species ),size=4)+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~TimePoint)+
    labs(y = "Biomass fraction per bin",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

  ggplot()+
    geom_point(data = hist.5.LogBinSummarize,aes(x=as.factor(LogBin),y=DensityLogBin,  color =Species ),size=4)+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~TimePoint)+
    labs(y = "Biomass fraction per bin",x = bquote(paste("Aggregate Volume (", mu*"m"^3*")")))+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))

#Separating aggregates vs planktonic
hist.plank.Agg.Sumary<-allDataHistograms.5df %>%  mutate(Aggregation = case_when(max == 5 ~ "Planktonic",max>5 ~ "Aggregate")) %>%
  ungroup() %>%group_by(Species, TimePoint, Condition,Aggregation) %>% 
  summarize(FrequencyAggregation = sum(frequency), BiomassAggregation = sum(BiomassBin), DensityAggregation = sum(Density))

hist.plank.Agg.Stats<-hist.plank.Agg.Sumary %>% group_by(Species,TimePoint, Condition,Aggregation) %>%
  summarise(MeanFreq = mean(FrequencyAggregation), SEFreq = std.error(FrequencyAggregation), MeanBiomass = mean(BiomassAggregation),
            SEBiomass = std.error(BiomassAggregation),MeanDensity = mean(DensityAggregation), SEDensity = std.error(DensityAggregation))

#Plotting these guys
PlotAggsPlank<- function(){
  ggplot()+
    geom_col(data = hist.plank.Agg.Stats,aes(x=TimePoint,y=MeanBiomass, fill = Aggregation),size=3,position = position_dodge( width = .9))+
    geom_errorbar(data = hist.plank.Agg.Stats,aes(x=TimePoint,ymin=MeanBiomass-SEBiomass, ymax = MeanBiomass+SEBiomass),size=.5,position = position_dodge2( width = .6))+
    facet_grid(.~Species)+
    labs(title = "Chronic Wounds", x = "", y = "Biomass contribution")+
    theme_cowplot()+
    theme(plot.title = element_text(hjust = 0.5))
}
PlotAggsPlank()
```

```{r biomass per slice}
#this was taken from "Microscopy data clean.Rmd"
GetResults<-function(sample,){#points to folder with files. Folder is hardcoded, unfortunately
  #folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Binary Images/', sample)
  folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Binary Images/wtd1-02')
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  blueChannel<-fileList[grep('B_',fileList)]
  for(slice in 1:length(filesList)/3){
    counts<-sum(as.array(readImage(fileList[1], convert = TRUE)))
    growthDf<-rbind(growthDf,c(counts, slice,species))
    growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
} }
#this was taken from the PropOc package
SlicesTo3D<-function(filesList,side){
  image1<-readTiff(filesList[1])#read first image to get its dimensions
  image3D <- array(0, c(image1@size[1], image1@size[2], length(filesList)))#preallocate array for image given using image dimensions
  for(i in 1:length(filesList)/3){#keep in mind that it is not zero-index
    #print(paste("loading ch1...", i, "/", length(filesList), filesList[i]))
    image3D[,,i]<-as.array(readImage(fileList[1], convert = TRUE))
  }
  return(image3D)
}

```

```{r Histogram functions}
FancyHistogram<-function(df, sec){#my own way of doing histograms which includes the sum and mean of values whitin each bin
  results<-data.frame()
  for( i in 1:(length(sec)-1)){ 
    blah<-df %>% filter(Size > sec[i] &Size<=sec[i+1])%>%summarise(frequency = n(), avg = mean(Size),suma = sum(Size)) %>% mutate(bin = i, min = sec[i],max = sec[i+1])
    if(blah$frequency>0){ 
      results<-rbind(results, blah)} 
  }
  return(results) 
}
TotalFancyHistogram<-function(df,binsize){
  totalbiomass<-sum(df$Size)#total biomass per sample
  totalfrequency<-length(df$Size)#total number of objects
  upperLim<-max(df$Size)+binsize-(max(df$Size)%%binsize)#largest bin size based in largest object
  breaksSequence<-seq(0,upperLim,binsize)#this makes breaks every "binsize" 
  histogram<-FancyHistogram(df,breaksSequence)#apply "fancy histogram" to get distributions
  histogram%>%mutate(BiomassBin=suma/totalbiomass)->histogram#add column with proportion of biomass per bin
  histogram%>%mutate(Density=frequency/totalfrequency)->histogram#add column with proportion of counts per bin
  histogram%>%mutate(Species=df$Species[1])->histogram#add column maintaining sample and so on...
  histogram%>%mutate(Condition=df$Condition[1])->histogram
  histogram%>%mutate(TimePoint=df$TimePoint[1])->histogram
  histogram%>%mutate(Sample=df$Sample[1])->histogram
  return(histogram)
}
```
